---
title: "Weight Lifting Exercise Dataset - prediction"
author: "Anna Witkowiak"
date: "26 10 2020"
output: html_document
---

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

## Loading data and libraries

```{r, echo=TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
training <- read.csv("~/Projekty/pml-training.csv", header = TRUE, na.strings = c('', 'NA'))
testing <- read.csv("~/Projekty/pml-testing.csv", header = TRUE, na.strings = c('', 'NA'))
```


## Cleaning data
The testing dataset has many variables that are mostly NA. I'm going to remove them.

```{r, echo=TRUE}
NAVAR    <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[, NAVAR==FALSE]
```



## Train and validation set
I need to split my training data into train and validation sets.
```{r, echo=TRUE}
set.seed(111)
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
ValSet  <- training[-inTrain, ]
```

## Models
Now it's time to build models and decide which one is better.

### 1. Decision tree
A Decision Tree is a supervised learning predictive model that uses a set of binary rules to calculate a target value.


```{r, echo=TRUE}
set.seed(222)
modDecTree <- train(classe ~ .,method="rpart",data=TrainSet,trControl=trainControl(number=5,method="cv",verboseIter=TRUE,allowParallel=TRUE))
print(modDecTree$finalModel)
plot(modDecTree$finalModel, uniform=TRUE, 
      main="Decision Tree")
text(modDecTree$finalModel, use.n=TRUE, all=TRUE, cex=.8)
```


```{r, echo=TRUE}
predDecTree <- predict(modDecTree, newdata=ValSet)
confDecTree <- confusionMatrix(predDecTree, ValSet$classe)
confDecTree
```

### 2. Random forest

Random forest consists of a large number of individual decision trees that operate as an ensemble. Each individual tree in the random forest spits out a class prediction and the class with the most votes becomes model’s prediction.


```{r, echo=TRUE}
set.seed(333)
modRForest <- train(classe ~ .,method="rf",data=TrainSet,trControl=trainControl(number=5,method="cv",verboseIter=TRUE,allowParallel=TRUE))
print(modRForest$finalModel)

```


```{r, echo=TRUE}
predRF <- predict(modRForest, newdata=ValSet)
confRF <- confusionMatrix(predRF, ValSet$classe)
confRF
```

## Selecting best model and predicting test data

Accuracy for Decision tree model is 0.66 and for Random forest model is 1.
This suggest that Random forest model might be overfitted. Even so I prefer to use that model over Decision tree.

```{r, echo=TRUE}
Testprediction <- predict(modRForest, newdata=testing)
Testprediction
```

